---
title: "P8130 Final Project - Group 19"
output: html_document
---

## Setup

##### General Setup
```{r general_setup}
knitr::opts_chunk$set(
  echo = TRUE,
  include = TRUE,
  message = FALSE,
  warning = FALSE
)
```

##### Visualizations
```{r visualiztion_setup}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

##### Installations
```{r installations}
library(tidyverse)
library(dplyr)
library(HH)
library(leaps)
library(corrplot) 
library(faraway)
library(broom)
```

##### Load Data Set
```{r}
hate_crime = read_csv("data/HateCrimes.csv", col_types="fffdddddd") %>% 
  janitor::clean_names()

head(hate_crime)
```

---

# Data Exploration: Descriptive statistics and visualization

Explore the distribution of the outcome and consider potential transformations if needed.(Vihar Desu)

##### Histogram of Outcome Variable

From the histogram below, we observe our outcome distribution has right skewness, suggesting that we may need to check our normality assumption. Our QQ Plot also indicates severe departures from normality.

```{r}
#Histogram of Outcome Distribution
hate_crime %>% 
  ggplot(aes(x = hate_crimes_per_100k_splc)) + 
  geom_histogram(color="red", fill="black") + 
  labs(
    title = "Distribution of Hate Crime Rates in 50 US States",
    x = "Hate Crime Rate per 100,000 Population",
    y = "Frequency of Distribution",
    caption = "Distribution of Hate Crime Rates ( 50 US States)")

#QQplot of Outcome Distribution
hate_crimes_per_100k_splc = hate_crime$hate_crimes_per_100k_splc
qqnorm(hate_crimes_per_100k_splc, col=2, pch=19, cex=1.5)
qq_plot = qqline(hate_crimes_per_100k_splc, col = 1,lwd=2,lty=2)

```

After performing a Shapiro-Wilk test to check the normality assumption of our outcome distribution, we find evidence to suggest that our data deviates from normality.

```{r}
# Perform Shapiro-Wilk test
shapiro.test(hate_crimes_per_100k_splc) %>% 
  broom::tidy() %>% 
  knitr::kable("simple")
```


Identify states with unusual rates and consider them as potential outliers/influential points. (Jyoti) \



# Income inequality was the main predictor of hate crime rates


Verify if this association holds true, as well as explore associations of all the other covariates mentioned above and draw your own conclusions.(Linh)\



Multicollinearity (Nikhita)\



Interaction between variables (e.g., urbanization), fit stratified models if the case. (Yanzhe Ma)\



# Model diagnostics 

Check model assumptions and goodness of fit (Vihar Desu)