---
title: "P8130 Final Project - Group 19"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(HH)
library(leaps)
library(corrplot) 
library(faraway)
library(broom)
```

Load the dataset

```{r}
hate_crime = read_csv("data/HateCrimes.csv")
```


# Data Exploration: Descriptive statistics and visualization


Explore the distribution of the outcome and consider potential transformations if needed.(Vihar Desu)\



Identify states with unusual rates and consider them as potential outliers/influential points. (Jyoti) \



# Income inequality was the main predictor of hate crime rates


Verify if this association holds true, as well as explore associations of all the other covariates mentioned above and draw your own conclusions.(Linh)\



Multicollinearity (Nikhita)\



Interaction between variables (e.g., urbanization), fit stratified models if the case. (Yanzhe Ma)\



# Model diagnostics 

Check model assumptions and goodness of fit (Vihar Desu)